### 2010-10-22 Thesis/Chapter7.tex
- Update 22 October -
Almost done...



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@188 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

An obvious modelling for thin structures is to apply the same methods than with solid objects. For instance, \cite{Holzapfel02} proposed a very complex 3-layer model of an aterial wall for simulating an angioplasty procedure. Their model is non-linear and anisotropic along fibers placed according to histology. They also model diseased part of the vessel through a hardening parameter which adds plasticity. However, they discretised their geometric model with eight-node brick elements. Another example is given by \cite{Aloisio04} who assume that the blood flow in the vessels may be considered as incompressible and the artery consequently becomes a solid body and can be modelled as a solid deformable object. If this approach can work, about a dozen of elements is required within the thickness of the structure for the system to be numerically stable \TODO{reference?}. Consequently, many elements are needed to model thin structures and real-time performance is difficult to achieve. Hence, specific models for thin objects were developed.

@@

Development of a satisfactory physical model that runs in real-time but produces visually convincing animation of thin objects has been a challenge in Computer Graphics, particularly in the area of cloth modelling. Rather than resorting to shell theory which relies on continuum mechanics, most of works have relied on discrete formulations. We will first discuss early approaches based on mass-spring models. Then we will present techniques relying on the derivation of a bending energy from geometric considerations. At last, we will introduce approaches based on shell theory using the more computationaly demanding finite element method.

@@

Early approaches to thin objects modelling only considered in-plane deformation, and often relied on mass-spring models. This technique was already presented in section \ref{chap4:massSprings} so we only present their use for modelling thin structures. \cite{Provot95} improved a mass-spring model to take into account the non-elastic properties of woven fabrics. Because of the high stiffness of textiles, mass-spring models are particulary unstable when used to simulate cloth. The author proposed a new method inspired from dynamic inverse procedure. Under certain conditions and because of the linear law employed in a mass-spring system, the fabrics may become super-elastic. \citeauthor{Provot95} suggested to set a threshold on the deformation rate to correct this super-elasticity (see \fig{chap7:fig-sheet}). 
%
\begin{figure}[ht]
\centering 
\subfloat[Initial position]{\includegraphics[width=4.5cm]{chapter7/sheet1.pdf}}
\hfill 
\subfloat[Uncorrected]{\includegraphics[width=4.5cm]{chapter7/sheet2.pdf}}
\hfill 
\subfloat[Corrected]{\includegraphics[width=4.5cm]{chapter7/sheet3.pdf}}
\caption{Deformation of a sheet hanging by two adjacent corners. Images courtesy of \cite{Provot95}.}
\label{chap7:fig-sheet}
\end{figure}

@@

A limitation of mass-spring models that we often hear is the difficulty to derive spring stiffness from elastic properties (Young's modulus and Poisson's ratio). With this drawback in mind, \cite{Volino97} improved a mass-spring system by allowing the modelisation of Young's modulus and Poisson's ratio (see \fig{chap7:fig-fabricSquares}). They also added a correction technique relying on the control of damping and velocities to ensure the stability of their model. 
%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=12cm]{chapter7/fabricSquares.pdf}
\caption[A $400\,$\% stretched fabric square]{A $400\,$\% stretched fabric square with various Poisson's ratio. Image courtesy of \cite{Volino97}.}
\label{chap7:fig-fabricSquares}
\end{center}
\end{figure}

Other works have considered adding bending through angular springs. 
%For instance, \cite{Taskiran05} have chosen to use a linear representation of angular springs to supply bending and torsion effects in hair modelling, which allowed them to simulate curly hair with good realism. All the forces related to springs were implemented: gravity, repulsions from collision, absorption, friction, etc. They could simulate up to a few hundreds of hair stripes in real-time. 
For instance, \cite{Wang07} successfully used a network of linear and angular springs to describe bending and twisting of catheters and guidewires in an interventional radiology simulator.

@@

\cite{Volino06} suggested an alternative somewhere between mass-spring and geometrically derived approaches. They compute a bending vector that represents the bending of the surface through  a simple linear combination of particule positions and they distribute this vector as particule forces according to the bending stiffness. The authors claimed much better accuracy than mass-spring models (particulary when dealing with low curvature and high bending stiffness) with similar computation time (see \fig{chap7:fig-dress} for an illustration). 
%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=10cm]{chapter7/dress.pdf}
\caption[Simulation of complete garments that have large bending stiffness]{The model of Volino and Magnenat-Thalmann allows the simulation of complete garments with material that may have fairly large bending stiffness. Image courtesy of \cite{Volino06}.}
\label{chap7:fig-dress}
\end{center}
\end{figure}

A method to use thin shell dynamics with point sampled surfaces for efficient animation was recently proposed by \cite{Wicke05} where the curvature of the shell is measured through the use of fibers. The fibers are used to approximate the differential surface operators. Their method supports both elastic and plastic deformations as well as fracturing and tearing of the material.

@@

\cite{Thomaszewski06} based their model on the Kirchhoff-Love shell theory and built a subdivision-based finite element formulation similar to \cite{Cirak00}.  They also employed a co-rotational framework to simulate wrinkles and folds of textiles (see \fig{chap7:fig-sleeve}). Although the rotation field is computed for each vertex, they found that using the rotation obtained for the barycentre of all vertices involved was sufficient. More recently, \cite{Allard09} also used a co-rotational finite element formulation to model the membrane of the lens capsule in a cataract surgery simulator. The goal was to simulate the anisotropic fracture progation during capsulorhexis, the technique used to create a circular opening in the lens capsule. The formulation did not feature bending however. 
%
\begin{figure}[ht]
\centering 
\subfloat[Under gravity]{\includegraphics[width=4.5cm]{chapter7/sleeve1.pdf}}
\hfill 
\subfloat[Torsion]{\includegraphics[width=4.5cm]{chapter7/sleeve2.pdf}}
\hfill 
\subfloat[Compression]{\includegraphics[width=4.5cm]{chapter7/sleeve3.pdf}}
\caption[Different types of folds on a garment's sleeve ]{Different types of folds on a garment's sleeve. Images courtesy of \cite{Thomaszewski06}.}
\label{chap7:fig-sleeve}
\end{figure}

>>>>

An obvious modelling for thin structures is to apply the same methods than with solid objects. For instance, \cite{Holzapfel02} proposed a very complex 3-layer model of an arterial wall for simulating an angioplasty procedure. Their model is non-linear and anisotropic along fibres placed according to histology. They also model diseased part of the vessel through a hardening parameter which adds plasticity. However, they discretised their geometric model with eight-node brick elements. Another example is given by \cite{Aloisio04} who assume that the blood flow in the vessels may be considered as incompressible and the artery consequently becomes a solid body and can be modelled as a solid deformable object. If this approach can work, about a dozen of elements is required within the thickness of the structure for the system to be numerically stable \TODO{reference?}. Consequently, many elements are needed to model thin structures and real-time performance is difficult to achieve. Hence, specific models for thin objects were developed.

@@

Development of a satisfactory physical model that runs in real-time but produces visually convincing animation of thin objects has been a challenge in Computer Graphics, particularly in the area of cloth modelling. Rather than resorting to shell theory which relies on continuum mechanics, most of works have relied on discrete formulations. We will first discuss early approaches based on mass-spring models. Then we will present techniques relying on the derivation of a bending energy from geometric considerations. At last, we will introduce approaches based on shell theory using the more computationally demanding finite element method.

@@

Early approaches to thin objects modelling only considered in-plane deformation, and often relied on mass-spring models. This technique was already presented in section \ref{chap4:massSprings} so we only present their use for modelling thin structures. \cite{Provot95} improved a mass-spring model to take into account the non-elastic properties of woven fabrics. Because of the high stiffness of textiles, mass-spring models are particularly unstable when used to simulate cloth. The author proposed a new method inspired from dynamic inverse procedure. Under certain conditions and because of the linear law employed in a mass-spring system, the fabrics may become super-elastic. \citeauthor{Provot95} suggested to set a threshold on the deformation rate to correct this super-elasticity (see \fig{chap7:fig-sheet}). 
%
\begin{figure}[ht]
\centering 
\subfloat[Initial position]{\includegraphics[width=4.5cm]{chapter7/sheet1.pdf}}
\hfill 
\subfloat[Uncorrected]{\includegraphics[width=4.5cm]{chapter7/sheet2.pdf}}
\hfill 
\subfloat[Corrected]{\includegraphics[width=4.5cm]{chapter7/sheet3.pdf}}
\caption{Deformation of a sheet hanging by two adjacent corners. Images courtesy of \cite{Provot95}.}
\label{chap7:fig-sheet}
\end{figure}

@@

A limitation of mass-spring models that we often hear is the difficulty to derive spring stiffness from elastic properties (Young's modulus and Poisson's ratio). With this drawback in mind, \cite{Volino97} improved a mass-spring system by allowing the modelling of Young's modulus and Poisson's ratio (see \fig{chap7:fig-fabricSquares}). They also added a correction technique relying on the control of damping and velocities to ensure the stability of their model. 
%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=12cm]{chapter7/fabricSquares.pdf}
\caption[A $400\,$\% stretched fabric square]{A $400\,$\% stretched fabric square with various Poisson's ratio. Image courtesy of \cite{Volino97}.}
\label{chap7:fig-fabricSquares}
\end{center}
\end{figure}

Other works have considered adding bending through angular springs. 
%For instance, \cite{Taskiran05} have chosen to use a linear representation of angular springs to supply bending and torsion effects in hair modelling, which allowed them to simulate curly hair with good realism. All the forces related to springs were implemented: gravity, repulsions from collision, absorption, friction, etc. They could simulate up to a few hundreds of hair stripes in real-time. 
For instance, \cite{Wang07} successfully used a network of linear and angular springs to describe bending and twisting of catheters and guide-wires in an interventional radiology simulator.

@@

\cite{Volino06} suggested an alternative somewhere between mass-spring and geometrically derived approaches. They compute a bending vector that represents the bending of the surface through  a simple linear combination of particle positions and they distribute this vector as particule forces according to the bending stiffness. The authors claimed much better accuracy than mass-spring models (particularly when dealing with low curvature and high bending stiffness) with similar computation time (see \fig{chap7:fig-dress} for an illustration). 
%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=10cm]{chapter7/dress.pdf}
\caption[Simulation of complete garments that have large bending stiffness]{The model of Volino and Magnenat-Thalmann allows the simulation of complete garments with material that may have fairly large bending stiffness. Image courtesy of \cite{Volino06}.}
\label{chap7:fig-dress}
\end{center}
\end{figure}

A method to use thin shell dynamics with point sampled surfaces for efficient animation was recently proposed by \cite{Wicke05} where the curvature of the shell is measured through the use of fibres. The fibres are used to approximate the differential surface operators. Their method supports both elastic and plastic deformations as well as fracturing and tearing of the material.

@@

\cite{Thomaszewski06} based their model on the Kirchhoff-Love shell theory and built a subdivision-based finite element formulation similar to \cite{Cirak00}.  They also employed a co-rotational framework to simulate wrinkles and folds of textiles (see \fig{chap7:fig-sleeve}). Although the rotation field is computed for each vertex, they found that using the rotation obtained for the barycentre of all vertices involved was sufficient. More recently, \cite{Allard09} also used a co-rotational finite element formulation to model the membrane of the lens capsule in a cataract surgery simulator. The goal was to simulate the anisotropic fracture propagation during capsulorhexis, the technique used to create a circular opening in the lens capsule. The formulation did not feature bending however. 
%
\begin{figure}[ht]
\centering 
\subfloat[Under gravity]{\includegraphics[width=4.5cm]{chapter7/sleeve1.pdf}}
\hfill 
\subfloat[Torsion]{\includegraphics[width=4.5cm]{chapter7/sleeve2.pdf}}
\hfill 
\subfloat[Compression]{\includegraphics[width=4.5cm]{chapter7/sleeve3.pdf}}
\caption[Different types of folds on a garment's sleeve ]{Different types of folds on a garment's sleeve. Images courtesy of \cite{Thomaszewski06}.}
\label{chap7:fig-sleeve}
\end{figure}

### 2010-10-22 Thesis/Chapter9.tex
- Update 22 October -
Almost done...



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@188 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{A co-rotational triangular shell FEM model}
\label{chap9}
\begin{shortAbstract}
In the field of medical simulation, very few models have been proposed for simulating, in real-time, the deformation of thin anatomical structures. In fact, to our knowledge, no model based on the equations of continuum mechanics was described and applied to simulating the deformation of objects. In this chapter, we propose to rely on continuum mechanics to accurately describe the complex deformations of the implant. Our approach extends the co-rotational method used in finite element analysis of in-plane deformations to incorporate a bending energy. We will first describe the model in details before discussing its implementation in the open source framework SOFA. Lastly, this co-rotational shell FEM was applied to the simulation of the lens deployment during catarct surgery. This simulation also accounts for the complex contacts that take place during the injection phase. This work was the object of two publications \citep{Comas2010a,Comas2010b}.
\end{shortAbstract}

@@

	\subsection{Contacts with the curved surface of shells}
If a specific rendering technique was developed to visualise the curved shape of the shells, it is also essential to interact with it. Indeed, as explained section \ref{chap9:interactions}, the force applied onto the curved surface of the shell must be processed by distributing the linear force received on the bent surface between the three vertices of the triangle. Consequently, the mapping previously described is extented to a mechanical mapping and two additional methods need to be implemented: applyJ and applyJT. 
%
\begin{itemize}
\item \textbf{applyJ()} encodes the linear relation between the DOF velocities and the shape velocities as described in section \ref{chap9:mappings}. applyJ() basically embodies the same computations than apply() except that the vector of displacement $\mathbf{u}$ is different. Whereas the local displacement in rotation is measured from the quaternions to create the vector $\mathbf{u}$, this time the vector is directly obtained from the angular velocities we had at each vertex (once converted into the local frame). We therefore have a relation between the angular velocities at each vertex and the velocity of the deflection in the direction normal to each triangle. 
\item \textbf{applyJT()} must convert the contact forces applied to the contact points to forces applied to the DOF. If the linear forces are easily distributed between DOFs using barycentric coordinates, it also encodes the relation between the vertical force $ \mathbf{F} $ applied onto the bent surface of the plate and the torque at each DOF. We start by retrieving the normal component of the applied force vector $F_{\mathrm{z}}$. We project the application point of the force into the triangle's plane and compute its local coordinates $(x,y)$. We create the polynom $P$ such as 
%
\begin{equation}
P = F_{\mathrm{z}}(1 \hspace{0.2cm} x \hspace{0.2cm} y \hspace{0.2cm} x^2 \hspace{0.2cm} xy \hspace{0.2cm} y^2 \hspace{0.2cm} x^3 \hspace{0.2cm} xy^2 \hspace{0.2cm} y^3)^T. 
\end{equation}
%
The moments at each vertex are then obtained with 
\begin{equation}
\Omega = (\textbf{C}^{-1})^T P.
\end{equation}
\end{itemize}

@@

To conclude on a possible parallelisation of the algorithm, an optimal GPU implementation is always a challenge to achieve. One needs to choose wisely the type of memory to be used for each variable and optimise the access patterns. However, it seems that there is no major obstacle to a GPU implementation and our co-rotation shell FEM will eventually be ported to GPU.

\TODO{enough details?}

\section{Comparison with analytical results}

@@

% Link where images from surgery have been found: http://asianophthalmology.org/articles/?a=articles&p=6
\begin{figure}[ht]
\centering
\includegraphics[width=4.5cm]{chapter9/surgery1.pdf}
\hfill
\includegraphics[width=4.5cm]{chapter9/surgery2.pdf}
\hfill
\includegraphics[width=4.5cm]{chapter9/surgery3.pdf} \\
\vspace{0.1cm}
\includegraphics[width=4.5cm]{chapter9/simu1.pdf}
\hfill
\includegraphics[width=4.5cm]{chapter9/simu2.pdf}
\hfill
\includegraphics[width=4.5cm]{chapter9/simu3.pdf}
\caption [Lens imlant] {Three steps of the simulation of the intra-ocular lens implant injection and its deployment within the lens capsule. Top: images from a real cataract surgery, courtesy of Dr. Tarek Youssef, ophtalmologist in Canada. Below: our simulation of the implant's deployment.}
\label{chap9:fig-simu-results}
\end{figure}

>>>>

\chapter{A co-rotational triangular shell FEM model}
\label{chap9}
\begin{shortAbstract}
In the field of medical simulation, very few models have been proposed for simulating, in real-time, the deformation of thin anatomical structures. In fact, to our knowledge, no model based on the equations of continuum mechanics was described and applied to simulating the deformation of objects. In this chapter, we propose to rely on continuum mechanics to accurately describe the complex deformations of the implant. Our approach extends the co-rotational method used in finite element analysis of in-plane deformations to incorporate a bending energy. We will first describe the model in details before discussing its implementation in the open source framework SOFA. Lastly, this co-rotational shell FEM was applied to the simulation of the lens deployment during cataract surgery. This simulation also accounts for the complex contacts that take place during the injection phase.
\end{shortAbstract}

@@

	\subsection{Contacts with the curved surface of shells}
If a specific rendering technique was developed to visualise the curved shape of the shells, it is also essential to interact with it. Indeed, as explained section \ref{chap9:interactions}, the force applied onto the curved surface of the shell must be processed by distributing the linear force received on the bent surface between the three vertices of the triangle. Consequently, the mapping previously described is extended to a mechanical mapping and two additional methods need to be implemented: applyJ and applyJT. 
%
\begin{itemize}
\item \textbf{applyJ()} encodes the linear relation between the DOF velocities and the shape velocities as described in section \ref{chap9:mappings}. applyJ() basically embodies the same computations than apply() except that the vector of displacement $\mathbf{u}$ is different. Whereas the local displacement in rotation is measured from the quaternions to create the vector $\mathbf{u}$, this time the vector is directly obtained from the angular velocities we had at each vertex (once converted into the local frame). We therefore have a relation between the angular velocities at each vertex and the velocity of the deflection in the direction normal to each triangle. 
\item \textbf{applyJT()} must convert the contact forces applied to the contact points to forces applied to the DOF. If the linear forces are easily distributed between DOFs using barycentric coordinates, it also encodes the relation between the vertical force $ \mathbf{F} $ applied onto the bent surface of the plate and the torque at each DOF. We start by retrieving the normal component of the applied force vector $F_{\mathrm{z}}$. We project the application point of the force into the triangle's plane and compute its local coordinates $(x,y)$. We create the polynom $P$ such as 
%
\begin{equation}
P = F_{\mathrm{z}}(1 \hspace{0.2cm} x \hspace{0.2cm} y \hspace{0.2cm} x^2 \hspace{0.2cm} xy \hspace{0.2cm} y^2 \hspace{0.2cm} x^3 \hspace{0.2cm} xy^2 \hspace{0.2cm} y^3)^T. 
\end{equation}
%
The moments at each vertex are then obtained with 
\begin{equation}
\Omega = (\textbf{C}^{-1})^T P.
\end{equation}
\end{itemize}

@@

Moreover, because shells are high-order (curved) elements, fewer elements are needed to describe a given geometry with the same precision (as compared to using flat triangles). Generally speaking, this constitutes a decisive computational advantage since fewer elements are required. However, GPUs are more efficient when they tackle numerous but simple computations. Hence, a good speedup over sequential CPU implementations is usually only achieved when the number of elements involved becomes great enough. In contrast, the computations for each shell element are quite heavy and the number of elements used in the whole simulation is limited. Therefore, the conditions are not optimal to obtain maximum speedups on GPU. A straightforward GPU implementation of our shell FEM formulation where each thread deals with the calculations of one element would not make use of all the GPU capacity (the occupancy would not be maximal). If a reasonable improvement over the CPU implementation will probably be obtained nevertheless, the gain is likely to be smaller than in the case of the TLED algorithm for instance. Other approaches will need to be investigated to take full advantage of the GPU. For instance, computations within an element could be distributed between several threads to increase the occupancy of the GPU and hence the overall efficiency.

To conclude on a possible parallelisation of the algorithm, an optimal GPU implementation is always a challenge to achieve. One needs to choose wisely the type of memory to be used for each variable and optimise the access patterns. While it seems that there is no major obstacle for a simple GPU implementation, the best speedups will only be attained through more elaborate techniques of parallelisation making full usage of the GPU. Although it will take some time, our co-rotation shell FEM will eventually be ported to GPU.

\section{Comparison with an analytical result}

@@

% Link where images from surgery have been found: http://asianophthalmology.org/articles/?a=articles&p=6
\begin{figure}[ht]
\centering
\includegraphics[width=4.5cm]{chapter9/surgery1.pdf}
\hfill
\includegraphics[width=4.5cm]{chapter9/surgery2.pdf}
\hfill
\includegraphics[width=4.5cm]{chapter9/surgery3.pdf} \\
\vspace{0.1cm}
\includegraphics[width=4.5cm]{chapter9/simu1.pdf}
\hfill
\includegraphics[width=4.5cm]{chapter9/simu2.pdf}
\hfill
\includegraphics[width=4.5cm]{chapter9/simu3.pdf}
\caption [Lens imlant] {Three steps of the simulation of the intra-ocular lens implant injection and its deployment within the lens capsule. Top: images from a real cataract surgery, courtesy of Dr. Tarek Youssef, ophthalmologist in Canada. Below: our simulation of the implant's deployment.}
\label{chap9:fig-simu-results}
\end{figure}

\section{Discussion}

The finite element procedure that we propose was largely developed based on physical understanding without the use of mathematical shell theories. Indeed, our shell FEM was obtained by simply superimposing a plane stress and bending energies. According to \cite{Chapelle03}, this type of approach has a limited level of accuracy. The effective study of shell structures clearly requires a deep physical understanding of shell structural behaviours. The complexity of the physical behaviours of shells requires advanced mathematical analyses from theories that are still in development. The interested reader may refer to the book of \cite{Chapelle03} for more details on shell theory. If the range of application of our shell formulation might be somewhat limited, we did not find out any inconsistencies through our tests. We also need to keep in mind that the accuracy required in medical simulation is not the same than the one demanded in structural analyses for instance. Moreover, the additional complexity of a shell finite element formulation based on true shell theory may also very well jeopardise our real-time constraints. Therefore, we do not think that the use of this simple shell formulation significantly diminishes our contribution to the field of medical simulation. Indeed, to our knowledge, our co-rotational shell FEM is the first description of a shell finite element model applied to simulating the deformation of thin anatomical structures.

Nevertheless, a few improvements are conceivable. Large strain could be allowed by using a non-linear strain tensor and more complex material laws (non-linear, viscoelasticity, etc.) could also be added to our shell formulation. However, those changes are substantial and their implementations are not straightforward. In addition, it is not clear how the benefits would be the greatest: whether from enhancing our model to a non-linear formulation or completely changing to another model based on true shell theory (where membrane and bending energies are not separated). The answer may depend on the type of constraints applied on the object. Also (as already mentioned section \ref{chap9:GPU}), the real-time capacity of our algorithm could be further improved by implementing our shell FEM on GPU. Although an optimal parallel re-implementation is always challenging, we have already reviewed why there was no major obstacle to a GPU implementation.

\section{Conclusion}

We proposed a co-rotational formulation for shell elements by extending a classical in-plane triangular finite element approach. This simple shell element can efficiently handle both membrane, bending and twist forces. The validity of our approach has been demonstrated though comparison with theoretical results. It was then applied it to a rather complex application: the simulation of intra-ocular lens implant deployment in a cataract surgery simulation. These results are very encouraging and show the potential of such models. Examples include the modelling of hollow anatomical structures (colon, stomach, etc.), the simulation of cardiac valve leaflets, and blood vessels. This work on this co-rotational shell FEM formulation was the object of one main publications \citep{Comas2010a}.

So far, the mesh of the implant used in the cataract surgery simulation was created manually. However, manually creating the mesh is not suitable if we want to model the deformation of complex anatomical structures using shell elements. Therefore, the first step is to describe the surface of the structure that we want to simulate with curved patches. The goal is to create a mesh featuring the optimal number of curved shell elements while staying as close as possible to our targeted geometry. The following chapter deals with this problem.

### 2010-10-22 Thesis/TitlePage.tex
- Update 11 October -
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@184 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b
- Update 12 October -
memoir class used now, fancy chapter style



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@185 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b
- Update 22 October -
Almost done...



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@188 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\begin{titlepage}

@@

\vspace*{2.5cm}
\Huge \textbf{Real-time Soft Tissue Modelling on GPU for Medical Simulation} \\
\vspace*{2.5cm}
\noindent \large Prepared by Olivier \textsc{Comas} at \\ 
INRIA Lille, \textsc{Shaman} team and CSIRO ICT Brisbane, AEHRC \\
\vspace*{0.5cm}
\noindent \large \color{red}{DRAFT - \today ~at \currenttime} \\
%\noindent \large Defended on the $00^{th}$ of December 2010 \\
\end{center}
\vspace*{1cm}
\noindent \large \textbf{Jury:} \\
\begin{center}
\noindent \large 
\begin{tabular}{llcl}
      \textit{Advisor:}	& Stéphane \textsc{Cotin}		& - & INRIA (\textsc{Shaman} team), Lille, France\\
      \textit{Reviewers:}	& Thomas Sangild \textsc{S\o rensen}          & - & University of Aarhus, Denmark\\
				& Georges \textsc{Dumont}		& - & ENS Cachan, Rennes, France\\
      \textit{President:}	& To be determined		& - & \\
      \textit{Examinators:}   & Tamy \textsc{Boubekeur}          & - & Telecom ParisTech, Paris, France\\
      				& Christian \textsc{Duriez}			& - & INRIA (\textsc{Shaman} team), Lille, France\\
      \textit{Invited:}   & Hans \textsc{De Visser}          & - & CSIRO, Brisbane, Australia\\
\end{tabular}
\end{center}
\end{titlepage}
\sloppy

\titlepage

>>>>

\thispagestyle{empty}

@@

\vspace*{2.5cm}
\Huge \textbf{Real-time Soft Tissue Modelling on GPU for Medical Simulation} \\
\vspace*{2.5cm}
\noindent \large Prepared by Olivier \textsc{Comas} at \\ 
INRIA Lille, \textsc{Shaman} team and CSIRO ICT Brisbane, AEHRC \\
\vspace*{0.5cm}
\noindent \large \color{red}{DRAFT - \today ~at \currenttime} \\
%\noindent \large Defended on the $00^{th}$ of December 2010 \\
\end{center}
\vspace*{1cm}
\noindent \large \textbf{Jury:} \\
\begin{center}
\noindent \large 
\begin{tabular}{llcl}
      \textit{Advisor:}	& Stéphane \textsc{Cotin}		& - & INRIA (\textsc{Shaman} team), Lille, France\\
      \textit{Reviewers:}	& Thomas Sangild \textsc{S\o rensen}          & - & University of Aarhus, Denmark\\
				& Georges \textsc{Dumont}		& - & ENS Cachan, Rennes, France\\
      \textit{Examiners:}   & Tamy \textsc{Boubekeur}          & - & Telecom ParisTech, Paris, France\\
      				& Christian \textsc{Duriez}			& - & INRIA (\textsc{Shaman} team), Lille, France\\
      \textit{President:}	& Laurent \textsc{Grisoni} & - & INRIA (\textsc{Mint} team), Lille, France \\
      \textit{Invited:}   & Hans \textsc{De Visser}          & - & CSIRO, Brisbane, Australia\\
\end{tabular}
\end{center}
\sloppy
